---
output:
  pdf_document:
    keep_tex: yes
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Evaluation against Douglas Peucker

Douglas Peucker is a wellknown algorithm \cite{dpeucker} in Computer Graphics and trajectory Computing. It can be implemented using a recursive approach. the algorithm reconstructs the trajectory by drawing  a line between the start and the end vertex and adding the one point on the trajectory that lies farthest away to split the line and recursively call itself two times once with the start point and the new point and once with the end point and the new point. It continus to recursively split until all points on the trajectory are closer to the new trajectory than the chosen value for epsilon

#Distance Measures

##discrete hausdorff distance

##Dynamic time warping distance

#Datasets

#Results

```{r, echo=FALSE}
library(trajcomp)
data(prague)

#get trajectory
prague$tid = getTrajectoryIDs(prague); # Add ID column
# and split on it
psplit = split (prague, f = prague$tid);


get_lengths <- function(fn, tr, ...){
  lens <-c()
  for(t in tr){
    x <- fn(as.matrix(t), ...)
    if(length(x)> 2)
      lens <- c(lens,length(x))
  }
  return(lens)
}

get_len_len <- function(tr){
  lens <-c()
  for(t in tr){
      lens <- c(lens, length(as.matrix(t)/3))
  }
  return(lens)
}
get_distances <- function(fn, tr, handle, ...){
  dists <-c()
  for(t in tr){
    x <- fn(as.matrix(t), ...)
    if(length(x)> 2){
      dists <- c(dists, distance_vector_ddply(as.matrix(x),as.matrix(t),handle))
    }
  }
  return(dists)
}


handle = compileSettings(
  tgroup(
    tdistance("discrete_hausdorff"),
    tdistance("dtw") #,  tdistance("discrete_frechet")
  ));

len_len <- get_len_len(psplit)
psp_len <- get_lengths(persistence_pruned, psplit, 16)
psm_len <- get_lengths(persistence_multires, psplit, 8, 7)
psd_len <- get_lengths(persistence_dist, psplit, 5,32, 2)
dp_len <- get_lengths(DouglasPeucker, psplit, 32)

psp_dst <- get_distances(persistence_pruned, psplit, handle, 16)
psm_dst <- get_distances(persistence_multires, psplit, handle, 8, 7)
psd_dst <- get_distances(persistence_dist, psplit, handle, 5,32, 2)
dp_dst <- get_distances(DouglasPeucker, psplit, handle, 32)
```

```{r, echo=FALSE}

plot(sort(dp_len), ylim=c(0,30), cex=1.7)
points(sort(psd_len), col="blue")
points(sort(psm_len), col="red", cex =0.5)
points(sort(psp_len), col="green", cex =0.2)
legend('topleft', 
       c("DP", "PS D", "PS MR", "PS PRU"), 
       col = c("black", "blue", "red", "green"), 
       lty = c(), 
       pch = c(1,1,1,1), 
       bty="n")

```

```{r, echo=FALSE}
hau <- c(TRUE, FALSE)
dtw <- c(FALSE, TRUE)

plot(sort(psm_dst[hau]), 
     ylim=c(0, 700), 
            cex=0.7, main="hausdorff distance",
     type="n")
lines(sort(dp_dst[hau]),  col="black")
lines(sort(psd_dst[hau]), col="blue")
lines(sort(psm_dst[hau]), col="red", cex =1.5)
lines(sort(psp_dst[hau]), col="green")
legend('topleft', 
       c("DP", "PS D", "PS MR", "PS PRU"), 
       col = c("black", "blue", "red", "green"), 
       lty = c(), 
       pch = c(1,1,1,1), 
       bty="n")
```

```{r, echo=FALSE}
plot(sort(dp_dst[dtw]),
     ylim=c(0, 7), 
     cex=0.7, main="DTW distance normalized", type="n")
lines(sort(dp_dst[dtw]/len_len), col="black")
lines(sort(psd_dst[dtw]/len_len), col="blue")
lines(sort(psm_dst[dtw]/len_len), col="red")
lines(sort(psp_dst[dtw]/len_len), col="green")
legend('topleft', 
       c("DP", "PS D", "PS MR", "PS PRU"), 
       col = c("black", "blue", "red", "green"), 
       lty = c(), 
       pch = c(1,1,1,1), 
       bty="n")
```

```{r, echo=FALSE}
plot(sort(dp_dst[dtw]),
     ylim=c(0, 3000), 
     cex=0.7, main="DTW distance", type="n")
lines(sort(dp_dst[dtw]), col="black")
lines(sort(psd_dst[dtw]), col="blue")
lines(sort(psm_dst[dtw]), col="red")
lines(sort(psp_dst[dtw]), col="green")
legend('topleft', 
       c("DP", "PS D", "PS MR", "PS PRU"), 
       col = c("black", "blue", "red", "green"), 
       lty = c(), 
       pch = c(1,1,1,1), 
       bty="n")
```

```{r, echo=FALSE}
plot(sort(dp_dst[hau]*psm_len),
     ylim=c(0, 10000), 
     cex=0.7, main="hausdorff * length", type="n")
lines(sort(dp_dst[hau]*dp_len), col="black")
lines(sort(psd_dst[hau]*psd_len), col="blue")
lines(sort(psm_dst[hau]*psm_len), col="red",)
lines(sort(psp_dst[hau]*psp_len), col="green")
legend('topleft', 
       c("DP", "PS D", "PS MR", "PS PRU"), 
       col = c("black", "blue", "red", "green"), 
       lty = c(), 
       pch = c(1,1,1,1), 
       bty="n")
```