---
title: "Persistence based trajectory simplification"
author: "Moritz Laass"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
Persistence based trajectory simplification is an algorithm that can be used to replace Douglas-Peucker for segmentation or simplification of trajectories. The main advantage of the algorithm is the complexity of `O(n)` compared to Douglas-Peucker's complexity of `O(n^2)`.

This package provides a few flavors of the the algorithm which shall be explored in the following sections, aswell as giving an understanding of the algorithms inner workings which will be usefull when deciding on the values to use for simplification or segmentation.


#Converting trajecories to curves
The persistence algorithm works on a curve that is derived from the discrete curvature in each point.
The package provides the function `persistence_curve` which converts a trajectory to a list of curvature values in degrees. It can be helpful to look at this curve if you are deciding on a value for `Beta` when using Beta-Pruning.

```{r, echo=FALSE}
library(trajcomp)
data(prague)

#get trajectory
prague$tid = getTrajectoryIDs(prague); # Add ID column
# and split on it
psplit = split (prague, f = prague$tid);
traj <- psplit[[1]];

#trajectory plot
plot(traj$x, traj$y, col="orange",
  main="Trajectory",
  xlab="x-axis",
  ylab="y-axis")
lines(traj$x, traj$y, col="black")

#curve plot
curve <- persistence_curve(as.matrix(traj))
plot(curve,  col="orange",
  main="Curve from Trajectory",
  xlab="Index",
  ylab="Curvature (Degrees)")
lines(curve, col="black")
```


The algorithm builds a connected component graph between the minima and maxima of the curve  which results in set of so called bars tha each connect a minima and a maxima.

By pruning the the bars, for which the difference between the minimum and the maximum is smaller than the pruning value `beta`, we get Beta-Pruning which will be discussed in the next section.


```{r, echo=FALSE}
library(trajcomp)
alpha  <- c(0, 9,0 ,-7,-2,-3,6, -2, 1,0)

plot(alpha, pch=1, col="orange", ylim = c(-7,22),
  main="Original Curve",
  xlab="Index",
  ylab="Curvature (Degrees)")

legend(1,25, c( "curve"), col = c("orange"), lty = c( 1), pch = c(1), bty="n")
lines(alpha, col="orange")

bars <- persistence_test_bars(alpha)
bars_min <- bars[c(TRUE, FALSE)]
bars_max <- bars[c(FALSE, TRUE)]

# curve and bar plot
plot(alpha, pch=1, col="orange", ylim = c(-7,22),
main="Curve with bars",
xlab="Index",
ylab="Curvature (Degrees)")

legend(5,25, c("bars", "curve"), col = c("blue", "orange"), lty = c(1, 1), pch = c(-1,1), bty="n")

legend(1,25, c("min", "max"), col = c("green", "red"), lty = c(-1, -1), pch = c(1,1), bty="n")
lines(alpha, col="orange")
points(bars_min+1, alpha[bars_min+1], col="green", pch=1)
points(bars_max+1, alpha[bars_max+1], col="red", pch=1)

xLow <- bars_min+1
yLow <- alpha[bars_min+1]
xHigh <- bars_max+1
yHigh <- alpha[bars_max+1]
arrows(xHigh,yHigh,xLow,yLow,col="blue",angle=90,length=0.1,code=3)

```


#Persistence with Beta-Pruning
Beta pruning removes all bars that fall below the threshhold of `beta`. For different types of trajectories different values of 'beta' yield better results. 

Also it should be noted that Beta pruning simplifies the trajectory well if  changes of direction happen gradually, but does not deal well with noisy data.

To deal with noise persistence with multiresolution simplification should be used, which will be discussed in the next section.

```{r, echo=FALSE}
library(trajcomp)
data(prague)

#get trajectory
prague$tid = getTrajectoryIDs(prague); # Add ID column
# and split on it
psplit = split (prague, f = prague$tid);
traj <- psplit[[1]];

#trajectory plot
plot(traj$x, traj$y, col="black", pch=21,
  main="Original Trajectory",
  xlab="x-axis",
  ylab="y-axis")
lines(traj$x, traj$y, col="orange")

pruned <- persistence_pruned(as.matrix(traj), 0)
#pruned trajectory plot
plot(traj$x[pruned+1], traj$y[pruned+1], col="black", pch=21,
  main="beta = 0",
  xlab="x-axis",
  ylab="y-axis")
lines(traj$x[pruned+1], traj$y[pruned+1], col="green")

pruned <- persistence_pruned(as.matrix(traj), 2)
plot(traj$x[pruned+1], traj$y[pruned+1], col="black", pch=21,
  main="beta = 2",
  xlab="x-axis",
  ylab="y-axis")
lines(traj$x[pruned+1], traj$y[pruned+1], col="blue")

pruned <- persistence_pruned(as.matrix(traj), 4)
plot(traj$x[pruned+1], traj$y[pruned+1], col="black", pch=21,
  main="beta = 4",
  xlab="x-axis",
  ylab="y-axis")
lines(traj$x[pruned+1], traj$y[pruned+1], col="red")

pruned <- persistence_pruned(as.matrix(traj), 8)
plot(traj$x[pruned+1], traj$y[pruned+1], col="black", pch=21,
  main="beta = 8",
  xlab="x-axis",
  ylab="y-axis")
lines(traj$x[pruned+1], traj$y[pruned+1], col="orange")


pruned <- persistence_pruned(as.matrix(traj), 16)
plot(traj$x[pruned+1], traj$y[pruned+1], col="black", pch=21,
  main="beta = 16",
  xlab="x-axis",
  ylab="y-axis")
lines(traj$x[pruned+1], traj$y[pruned+1], col="purple")

```

#Persistence with Multi Resolution Simplification

Multi resolution takes a second parameter, additional to `beta`, called `levels`. 
Levels is an integer value that defines the number of iterations the trajectoriy goes through beta pruning and a noise reduction step which removes all vertices that are closer  to each other than '2^(iteration)' . That means for trajectories, that cover bigger distances larger values of `levels` should be used. 

Also it is noteworthy to select both `beta` and `levels` accorinf to your data, to get ideal results.

##Multiresolution Simplification with Beta = 2
```{r, echo=FALSE}
library(trajcomp)
data(prague)

#get trajectory
prague$tid = getTrajectoryIDs(prague); # Add ID column
# and split on it
psplit = split (prague, f = prague$tid);
traj <- psplit[[1]];


#trajectory plot
plot(traj$x, traj$y, col="black", pch=21,
  main="original trajectory",
  xlab="x-axis",
  ylab="y-axis")
lines(traj$x, traj$y, col="orange")


pruned <- persistence_multires(as.matrix(traj), 2, 1)
plot(traj$x[pruned+1], traj$y[pruned+1], col="black", pch=21,
  main="beta = 2, levels = 1",
  xlab="x-axis",
  ylab="y-axis")
lines(traj$x[pruned+1], traj$y[pruned+1], col="green")

pruned <- persistence_multires(as.matrix(traj), 2, 2)
plot(traj$x[pruned+1], traj$y[pruned+1], col="black", pch=21,
  main="beta = 2, levels = 2",
  xlab="x-axis",
  ylab="y-axis")
lines(traj$x[pruned+1], traj$y[pruned+1], col="blue")

pruned <- persistence_multires(as.matrix(traj), 2, 3)
plot(traj$x[pruned+1], traj$y[pruned+1], col="black", pch=21,
  main="beta = 2, levels = 3",
  xlab="x-axis",
  ylab="y-axis")
lines(traj$x[pruned+1], traj$y[pruned+1], col="red")

pruned <- persistence_multires(as.matrix(traj), 2, 4)
plot(traj$x[pruned+1], traj$y[pruned+1], col="black", pch=21,
  main="beta = 2, levels = 4",
  xlab="x-axis",
  ylab="y-axis")
lines(traj$x[pruned+1], traj$y[pruned+1], col="orange")

pruned <- persistence_multires(as.matrix(traj), 2, 5)
plot(traj$x[pruned+1], traj$y[pruned+1], col="black", pch=21,
  main="beta = 2, levels = 5",
  xlab="x-axis",
  ylab="y-axis")
lines(traj$x[pruned+1], traj$y[pruned+1], col="purple")

```


##Multiresolution Simplification with Beta = 4
```{r, echo=FALSE}
library(trajcomp)
data(prague)

#get trajectory
prague$tid = getTrajectoryIDs(prague); # Add ID column
# and split on it
psplit = split (prague, f = prague$tid);
traj <- psplit[[1]];


#trajectory plot
plot(traj$x, traj$y, col="black", pch=21,
  main="original trajectory",
  xlab="x-axis",
  ylab="y-axis")
lines(traj$x, traj$y, col="orange")


pruned <- persistence_multires(as.matrix(traj), 4, 1)
plot(traj$x[pruned+1], traj$y[pruned+1], col="black", pch=21,
  main="beta = 4, levels = 1",
  xlab="x-axis",
  ylab="y-axis")
lines(traj$x[pruned+1], traj$y[pruned+1], col="green")

pruned <- persistence_multires(as.matrix(traj), 4, 2)
plot(traj$x[pruned+1], traj$y[pruned+1], col="black", pch=21,
  main="beta = 4, levels = 2",
  xlab="x-axis",
  ylab="y-axis")
lines(traj$x[pruned+1], traj$y[pruned+1], col="blue")

pruned <- persistence_multires(as.matrix(traj), 4, 3)
plot(traj$x[pruned+1], traj$y[pruned+1], col="black", pch=21,
  main="beta = 4, levels = 3",
  xlab="x-axis",
  ylab="y-axis")
lines(traj$x[pruned+1], traj$y[pruned+1], col="red")

pruned <- persistence_multires(as.matrix(traj), 4, 4)
plot(traj$x[pruned+1], traj$y[pruned+1], col="black", pch=21,
  main="beta = 4, levels = 4",
  xlab="x-axis",
  ylab="y-axis")
lines(traj$x[pruned+1], traj$y[pruned+1], col="orange")

pruned <- persistence_multires(as.matrix(traj), 4, 5)
plot(traj$x[pruned+1], traj$y[pruned+1], col="black", pch=21,
  main="beta = 4, levels = 5",
  xlab="x-axis",
  ylab="y-axis")
lines(traj$x[pruned+1], traj$y[pruned+1], col="purple")

```



##Multiresolution Simplification with Beta = 8
```{r, echo=FALSE}
library(trajcomp)
data(prague)

#get trajectory
prague$tid = getTrajectoryIDs(prague); # Add ID column
# and split on it
psplit = split (prague, f = prague$tid);
traj <- psplit[[1]];


#trajectory plot
plot(traj$x, traj$y, col="black", pch=21,
  main="original trajectory",
  xlab="x-axis",
  ylab="y-axis")
lines(traj$x, traj$y, col="orange")


pruned <- persistence_multires(as.matrix(traj), 8, 1)
plot(traj$x[pruned+1], traj$y[pruned+1], col="black", pch=21,
  main="beta = 8, levels = 1",
  xlab="x-axis",
  ylab="y-axis")
lines(traj$x[pruned+1], traj$y[pruned+1], col="green")

pruned <- persistence_multires(as.matrix(traj), 8, 2)
plot(traj$x[pruned+1], traj$y[pruned+1], col="black", pch=21,
  main="beta = 8, levels = 2",
  xlab="x-axis",
  ylab="y-axis")
lines(traj$x[pruned+1], traj$y[pruned+1], col="blue")

pruned <- persistence_multires(as.matrix(traj), 8, 3)
plot(traj$x[pruned+1], traj$y[pruned+1], col="black", pch=21,
  main="beta = 8, levels = 3",
  xlab="x-axis",
  ylab="y-axis")
lines(traj$x[pruned+1], traj$y[pruned+1], col="red")

pruned <- persistence_multires(as.matrix(traj), 8, 4)
plot(traj$x[pruned+1], traj$y[pruned+1], col="black", pch=21,
  main="beta = 8, levels = 4",
  xlab="x-axis",
  ylab="y-axis")
lines(traj$x[pruned+1], traj$y[pruned+1], col="orange")

pruned <- persistence_multires(as.matrix(traj), 8, 5)
plot(traj$x[pruned+1], traj$y[pruned+1], col="black", pch=21,
  main="beta = 8, levels = 5",
  xlab="x-axis",
  ylab="y-axis")
lines(traj$x[pruned+1], traj$y[pruned+1], col="purple")

```

