---
title: "TRACLUS trajectory clustering"
author: "Martin Werner"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
TRACLUS is a framework for trajectory clustering building on density-based DBSCAN algorithm together with
segmentation and grouping functionalities. Basically, all trajectories are split into reasonably short segments,
then these segments are clustered, clusters of trajectories are extracted from clusters of segments and
the final result is returned. 

# Basic Example

Let us first load the library and a dataset (smallprague) and apply traclus.

```{r}
library(trajcomp)
data(smallprague)
segments = traclus(smallprague, 25,5)
head(segments)
```

This returns all segments as created by the internal segmentation of TRACLUS algorithm
together with a cluster ID. Negative clusters are noise.

```{r}
library(plyr) # for efficient grouping using ddply

# plot the dataset into the background
plot(smallprague,t="l",col="gray")

# not actually elegant, but very clear for beginners:
cluster_to_color <- function(r)
{
   if (r == -2){
      return ("red");
   }
   if (r == -1){
       return("black");
   }
   if (r == 0){
       return ("blue");
   }
   if(r == 1){
      return ("green");
   }
}
# Group by cluster and draw color

ddply(segments,.(cluster),function(x){
   points(c(x$x1,x$x2),c(x$y1,x$y2),col=cluster_to_color(x$cluster[1]));
})
```
